---
title: "Capstone Project:\nAnalysing Dengue Cases in Singapore"
author: "Shayne, Roscoe, Anu"
date: "05 July 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

library(ggplot2);library(magrittr)
```

```{r, import, include=FALSE}
get_moh_eweekly <- function(path) {
  #' Get (Epi) Weekly Records from Ministry of Health (MOH)
  #' 
  #' @description
  #' Available data ranges from 2012-W01 to 2020-W25. The data file is an Excel Workbook with multiple sheets. The file has to be downloaded first. Each sheet stores a year's data.
  #' 
  #' @details https://www.moh.gov.sg/docs/librariesprovider5/diseases-updates/weekly-infectious-disease-bulletin-year-2020ea2c0b1cec1549009844537d52f2377f.xlsx
  #' 
  #' @param path The file path of the dataset.
  #' @return A table containing the combined historical epi-weekly records.
  #' @examples
  #' get_moh_eweekly("dataset.xlsx")
  
  # Read multiple sheets in an Excel Workbook
  dfs = path %>% 
    readxl::excel_sheets() %>% 
    set_names(., .) %>%
    lapply(readxl::read_xlsx, path = path, skip = 1)
  
  # Start and End date formats are different for 2020
  dfs[["2020"]]$Start = lubridate::dmy(dfs[["2020"]]$Start)
  dfs[["2020"]]$End = lubridate::dmy(dfs[["2020"]]$End)
  
  for (y in names(dfs)) {
    # Add a "Year" column to distinguish different tables
    dfs[[y]]$Year = y
    
    # Standardize naming of column headers
    names(dfs[[y]])[names(dfs[[y]]) %in% c("Campylobacter enterosis",
                                           "Campylobacterenterosis",
                                           "Campylobacteriosis")] <- "Campylobacter enteritis"
    names(dfs[[y]])[names(dfs[[y]]) == "Chikungunya Fever"] <- "Chikungunya"
    names(dfs[[y]])[names(dfs[[y]]) == "Dengue Haemorrhagic Fever"] <- "DHF"
    names(dfs[[y]])[names(dfs[[y]]) == "Dengue Fever"] <- "Dengue"
    names(dfs[[y]])[names(dfs[[y]]) %in% c("Hand, Foot and Mouth Disease",
                                           "Hand, Foot Mouth Disease")] <- "HFMD"
    names(dfs[[y]])[names(dfs[[y]]) == "Nipah virus infection"] <- "Nipah"
    names(dfs[[y]])[names(dfs[[y]]) == "Viral Hepatitis A"] <- "Acute Viral Hepatitis A"
    names(dfs[[y]])[names(dfs[[y]]) == "Viral Hepatitis E"] <- "Acute Viral Hepatitis E"
    names(dfs[[y]])[names(dfs[[y]]) %in% c("Zika Virus Infection",
                                           "Zika virus infection")] <- "Zika"
  }
  
  dplyr::bind_rows(dfs) %>% 
    dplyr::select(Year, everything()) %>% 
    dplyr::arrange(Year, `Epidemiology Wk`) %>% 
    dplyr::mutate(Year = as.numeric(Year))
}

get_mss_daily <- function(years, stations = c("Changi")) {
  #' Get Daily Records from Meteorological Service Singapore (MSS)
  #' 
  #' @description
  #' Available data ranges from Jan 1980 to May 2020. Epidemiological weeks will be calculated automatically.
  #' 
  #' @details http://www.weather.gov.sg/climate-historical-daily/
  #' 
  #' @param years A vector of years of interest.
  #' @param stations A vector of climate station names. Defaults to c("Changi").
  #' @return A table containing the combined historical daily records.
  #' @examples
  #' get_mss_daily(2014:2018)
  #' get_mss_daily(2012:2020, c("Changi", "Marine Parade", "Queenstown", "Sembawang"))
  
  # All combinations of the given years and the 12 months
  dates = as.vector(t(outer(years, sprintf("%02d", 1:12), FUN = paste0)))
  
  # Remove the last 7 months if 2020 is included (no data after May 2020)
  if (2020 %in% years) {
    dates = dates[1:(length(dates) - 7)]
  }
  
  mappings = list(
    "Admiralty" = "104_",
    "Ang Mo Kio" = "109_",
    "Changi" = "24_",
    "Choa Chu Kang (South)" = "121_",
    "Clementi" = "50_",
    "East Coast Parkway" = "107_",
    "Jurong (West)" = "44_",
    "Jurong Island" = "117_",
    "Khatib" = "122_",
    "Marina Barrage" = "108_",
    "Newton" = "111_",
    "Pasir Panjang" = "116_",
    "Pulau Ubin" = "106_",
    "Seletar" = "25_",
    "Sembawang" = "80_",
    "Sentosa Island" = "60_",
    "Tai Seng" = "43_",
    "Tengah" = "23_",
    "Tuas South" = "115_"
    
    # "Boon Lay (East)" = "86_",
    # "Buangkok" = "55_",
    # "Bukit Panjang" = "64_",
    # "Lower Peirce Reservoir" = "08_",
    # "Marine Parade" = "113_",
    # "Punggol" = "81_",
    # "Queenstown" = "77_",
    # "Semakau Island" = "102_",
    # "Serangoon" = "36_",
    # "Toa Payoh" = "88_",
    # "Yishun" = "91_"
  )
  
  station_nums = sapply(stations, function(x) mappings[[x]])
  urls = paste0("http://www.weather.gov.sg/files/dailydata/DAILYDATA_S", 
                t(outer(station_nums, dates, FUN = paste0)), 
                ".csv")
  
  myreader = function(url) {
    tryCatch(
      readr::read_csv(url) %>% 
        # The "degree sign" in column headers are invalid multibyte characters
        # Data after Apr 2020 have slightly different column names
        # This would complicate joining the tables together
        # So, it might be better to manually label all column names
        setNames(., c("Station",
                      "Year",
                      "Month",
                      "Day",
                      "Daily_Rainfall_Total_mm",
                      "Highest_30_Min_Rainfall_mm",
                      "Highest_60_Min_Rainfall_mm",
                      "Highest_120_Min_Rainfall_mm",
                      "Mean_Temperature_degC",
                      "Maximum_Temperature_degC",
                      "Minimum_Temperature_degC",
                      "Mean_Wind_Speed_kmh",
                      "Max_Wind_Speed_kmh")) %>% 
        # The "long dashes" (em dash?) representing missing values are invalid
        #   multibyte characters and are replaced with NA (by coercion)
        dplyr::mutate_at(dplyr::vars(-Station), as.numeric),
      error = function(e) { NA }
    )
  }
  
  dfs = lapply(urls, myreader)
  
  dplyr::bind_rows(dfs[!is.na(dfs)]) %>% 
    # Calculate epidemiological weeks (for joining with other datasets)
    dplyr::mutate(Date = paste(Year, Month, Day, sep = "-"),
                  Epiweek = lubridate::epiweek(Date)) %>% 
    dplyr::select(Station, Epiweek, everything(), -Date)
}

# w_data <- get_mss_daily(years = 2012:2020,
#                     stations = c("Changi",
#                                  "Marine Parade",
#                                  "Queenstown",
#                                  "Sembawang"))

# w_data <- get_mss_daily(years = 2012:2020,
#                     stations = c("Admiralty",
#                                  "Ang Mo Kio",
#                                  "Changi",
#                                  "Choa Chu Kang (South)",
#                                  "Clementi",
#                                  "East Coast Parkway",
#                                  "Jurong (West)",
#                                  "Jurong Island",
#                                  "Khatib",
#                                  "Marina Barrage",
#                                  "Newton",
#                                  "Pasir Panjang",
#                                  "Pulau Ubin",
#                                  "Seletar",
#                                  "Sembawang",
#                                  "Sentosa Island",
#                                  "Tai Seng",
#                                  "Tengah",
#                                  "Tuas South"))

weather <- readr::read_csv("../results/weather_daily_2012_2020.csv")
bulletin <- get_moh_eweekly("../data/weekly-infectious-disease-bulletin-year-2020ea2c0b1cec1549009844537d52f2377f.xlsx")
```

```{r, transform, include=FALSE}
grouped_epiweek_changi <- weather %>%
  dplyr::filter(Station == "Changi") %>%
  dplyr::select(Year, Epiweek, matches("Rainfall_Total|Temperature"), -Station) %>% 
  dplyr::group_by(Year, Epiweek) %>% 
  dplyr::summarise(Rainfall = sum(Daily_Rainfall_Total_mm),
                   Temperature = mean(Mean_Temperature_degC))

df <- bulletin %>% 
  dplyr::rename(Epiweek = `Epidemiology Wk`) %>% 
  dplyr::select(Year, Epiweek, Start, Dengue) %>% 
  tidyr::drop_na() %>% 
  dplyr::left_join(grouped_epiweek_changi, c("Year", "Epiweek"))
```

---

```{r, viz_ncases, echo=FALSE, fig.width=8, fig.height=6}
f1 <- df %>% 
  ggplot(aes(x = Start, y = Dengue)) + 
  geom_line(size = 1, color = "tomato3", alpha = 0.7) + 
  labs(title = "Weekly Dengue Cases from 2012 to 2020",
       subtitle = "Why is there a sudden spike in dengue cases this year?",
       x = "",
       y = "Number of Cases",
       caption = "Source: moh.gov.sg")

f1
```



## Overview of the Project {data-background=https://research.a-star.edu.sg/cms/figure/index/55e43a55140ba0db748b4b69.jpg data-background-size="20% 20%" data-background-position="right top"}
Dengue fever is a vector-borne infectious disease that are endemic in the tropical world. Singapore is one of several countries with high disease burden of dengue. In 2020, Singapore saw 1,158 dengue cases in a week of June - the highest number of weekly dengue cases ever recorded since 2014. Why is there a sudden spike in dengue cases this year?

### Questions
  - Does atmospheric variables influence the incidence of dengue cases?
    - Specifically, is higher humidity, precipitation, or temperature associated with increased numbers of dengue cases?
  - Can atmospheric variables predict the number of dengue cases?
  - Does the number of dengue cases increase with the number of COVID-19 cases?
  - Explain the different number of cases in different regions of Singapore



## Data
- Daily climate records from [Meteorological Service Singapore (MSS)](http://www.weather.gov.sg/climate-historical-daily/)
  - Relevant variables: Rainfall, Temperature
  - Data for 2012 to 2020 will be gathered from selected climate stations
- Weekly case numbers (2012-W01 to 2020-W25) from [Ministry of Health (MOH)](https://www.moh.gov.sg/docs/librariesprovider5/diseases-updates/weekly-infectious-disease-bulletin-year-2020ea2c0b1cec1549009844537d52f2377f.xlsx)
- Latest number of cases by named regions in Singapore from [National Environment Agency (NEA)](https://www.nea.gov.sg/dengue-zika/dengue/dengue-clusters)
- Yearly population distribution across named regions in Singapore
- COVID-19 cases (Apr - Jul 2020)

## Deprecated Data
- Monthly Air Temperature And Sunshine, Relative Humidity And Rainfall from [Singapore Department of Statistics (DOS)](https://www.tablebuilder.singstat.gov.sg/publicfacing/api/csv/title/15306.csv)
  - <u>Reason</u>: Higher resolution (daily) data available from MSS
  - Might reconsider if humidity data is needed
- Latest number of cases by regions with geocoordinates from [Data.gov.sg](https://data.gov.sg/dataset?q=Dengue)
  - <u>Reason</u>: Potentially **grandiose** (we've not learned leaflet yet!)
- Weekly case numbers (2012-W01 to 2020-W20) from [Data.gov.sg](https://data.gov.sg/dataset?q=Dengue)
  - <u>Reason</u>: More up-to-date data available from MOH



## Analysis Plan
- Transform dengue cases data
  - Converting epidemiology weeks cases to monthly cases for 2012 - 2020
  - Combine the different Excel sheets into a table 
- Compare the monthly cases across the years
  - Using repeated-measures ANOVA
  - Plotting
- To understand if atmospheric variables have an influence in the incidence of dengue cases
  - Correlate the weather variables with number of dengue cases monthly across the years
- To test if atmospheric variables predict number of dengue cases
  - Create models for regression
  - Compare the predicted value with actual value

## Analysis Plan
- For the current trend, explain the differences between the number of cases in different regions
  - According to the news report, East region has the largest number of clusters compared to other parts of Singapore
  - In particular, does Marine Parade and Sembawang (or Queenstown?) have any differences in their weather variables?

### Considerations and limitations
- Time lag: Dengue cases manifest 1-2 weeks after infection
  - Timing adjustments have to be made for ostensibly associated variables
- Seasonal effects
  - "Vector" months (June, July, August, September, October)
- COVID-19 cases vs. dengue cases
  - 90% foreign workers, expect a correlation
  - Compare against number of cases from Apr - Jul 2020



# Preliminary Survey of Data Collected so Far

---

```{r, viz_grid, echo=FALSE, warning=FALSE, fig.width=8, fig.height=6}
f2 <- df %>% 
  ggplot(aes(x = Start, y = Rainfall)) + 
  geom_line(size = 1, color = "deepskyblue4", alpha = 0.7) + 
  labs(title = "Total weekly Rainfall from 2012 to 2020",
       subtitle = "",
       x = "",
       y = "Total Rainfall (mm)",
       caption = "Source: moh.gov.sg")

f3 <- df %>% 
  ggplot(aes(x = Start, y = Temperature)) + 
  geom_line(size = 1, color = "darkgreen", alpha = 0.7) + 
  labs(title = "Average weekly Temperature from 2012 to 2020",
       subtitle = "",
       x = "",
       y = "Temperature (\u00b0C)",
       caption = "Source: moh.gov.sg")

gridExtra::grid.arrange(f1, f2, f3, nrow = 3, top = "Is there a relationship?")
```

---

```{r pre_explore, include=FALSE}
df <- bulletin %>% 
  dplyr::rename(Epiweek = `Epidemiology Wk`) %>% 
  dplyr::select(Year, Epiweek, Start, Dengue) %>% 
  tidyr::drop_na() %>% 
  dplyr::left_join(grouped_epiweek_changi, c("Year", "Epiweek"))

grouped_month <- weather %>%
  dplyr::select(Station, Year, Month, matches("Rainfall_Total|Temperature")) %>% 
  dplyr::group_by(Station, Year, Month) %>% 
  dplyr::summarise(Rainfall_Total = sum(Daily_Rainfall_Total_mm),
                   Mean_Temperature = mean(Mean_Temperature_degC),
                   Maximum_Temperature = mean(Maximum_Temperature_degC),
                   Minimum_Temperature = mean(Minimum_Temperature_degC))

station_positions <- readr::read_csv("../data/Station_Records.csv") %>% 
  dplyr::rename(Lat = `Lat.(N)`,
                Long = `Long. (E)`) %>% 
  dplyr::select(Station,
                Lat,
                Long,
                `Period of Daily Rain Records`, 
                `Period of Mean Temperature`,
                `Period of Max and Min Temperature`) %>% 
  tidyr::drop_na() %>% 
  dplyr::select(Station, Lat, Long) %>%
  dplyr::arrange(Station)



lspdf <- rgdal::readOGR("../data/master-plan-2019-planning-area-boundary-no-sea/planning-boundary-area.kml")

lspdf@data$area_name <- gsub(".*?<td>(.*?)</td>.*", "\\1", lspdf@data$Description)
centers <- rgeos::gCentroid(lspdf, byid = T)
centers$region <- lspdf@data$area_name

pal <- leaflet::colorFactor(RColorBrewer::brewer.pal(nrow(lspdf@data), "Set3"), NULL)
```

## Climate Stations

```{r, viz_stations, echo=FALSE}
station_positions %>%
  leaflet::leaflet(width = "100%") %>% 
  leaflet::addTiles() %>% 
  leaflet::addMarkers(~Long, 
                      ~Lat, 
                      popup = ~as.character(Station), 
                      label = ~as.character(Station), 
                      labelOptions = leaflet::labelOptions(noHide = T))
```

## Climate Stations with Planning Areas Overlay

```{r, viz_planning_areas, echo=FALSE}
lspdf %>% 
  leaflet::leaflet(width = "100%") %>%
  leaflet::addTiles() %>%
  leaflet::addPolygons(stroke = T,
                       opacity = 1,
                       # color = "black",
                       smoothFactor = 0.5,
                       fillOpacity = 0.75,
                       fillColor = ~pal(area_name),
                       weight = 1,
                       label = lspdf@data$area_name,
                       labelOptions = leaflet::labelOptions(
                         style = list("font-weight" = "normal", padding = "3px 8px"),
                         textsize = "15px",
                         direction = "auto",
                         noHide = F
                       )) %>% 
  leaflet::addLabelOnlyMarkers(data = centers,
                               lng = centers$x,
                               lat = centers$y,
                               label = ~region,
                               labelOptions = leaflet::labelOptions(noHide = T,
                                                           direction = "top",
                                                           textOnly = T)) %>% 
  leaflet::addMarkers(data = station_positions,
                      lng = ~Long,
                      lat = ~Lat,
                      popup = ~as.character(Station),
                      label = ~as.character(Station),
                      labelOptions = leaflet::labelOptions(noHide = F))
```

---

```{r, explore_weather, echo=FALSE, fig.width=8, fig.height=5}
w1 <- grouped_month %>% 
  dplyr::filter(Station == "Changi") %>% 
  tibble::add_column(x = 1:nrow(.)) %>% 
  dplyr::select(-Rainfall_Total) %>% 
  tidyr::gather("Variable",
                "Value",
                Mean_Temperature,
                Maximum_Temperature,
                Minimum_Temperature) %>% 
  ggplot(aes(x = x, y = Value, color = Variable)) + 
  scale_x_continuous(breaks = c(1 + 12 * (0:8)),
                     label = c(2012:2020)) +
  geom_line(size = 1) + 
  labs(title = "Average monthly Temperature from 2012 to 2020",
       subtitle = "",
       x = "",
       y = "Temperature (\u00b0C)",
       caption = "Source: weather.gov.sg")

w2 <- grouped_month %>% 
  dplyr::select(Station, Year, Month, Rainfall_Total) %>% 
  dplyr::mutate(x = (Year - 2012) * 12 + Month) %>% 
  ggplot(aes(x = x, y = Rainfall_Total, color = Station)) + 
  scale_x_continuous(breaks = c(1 + 12 * (0:8)),
                     label = c(2012:2020)) +
  geom_line(size = 1, alpha = 0.5) + 
  labs(title = "Total monthly Rainfall from 2012 to 2020",
       subtitle = "",
       x = "",
       y = "Total Rainfall (mm)",
       caption = "Source: weather.gov.sg")

gridExtra::grid.arrange(w1, w2, nrow = 2)
```

---

```{r, viz_density, echo=FALSE, warning=FALSE, fig.width=8, fig.height=3}
{
  par(mfrow = c(1, 3))
  plot(density(df$Dengue[!is.na(df$Dengue)]), main = "Cases")
  plot(density(df$Rainfall[!is.na(df$Rainfall)]), main = "Rainfall")
  plot(density(df$Temperature[!is.na(df$Temperature)]), main = "Temperature")
}
```
