---
title: "Capstone Project Proposal:\nAnalysing Dengue Cases in Singapore"
author: "S, R, A"
date: "04 July 2020"
output:
  ioslides_presentation:
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

library(ggplot2);library(magrittr)
```

---

```{r, import, fig.cap="", echo=FALSE, include=FALSE}
get_moh_eweekly <- function(path) {
  #' Get (Epi) Weekly Records from Ministry of Health (MOH)
  #' 
  #' @description
  #' Available data ranges from 2012-W01 to 2020-W25.
  #' 
  #' The file is a multi-sheet Excel Workbook. Each sheet stores a year's data.
  #' 
  #' @details
  #' "https://www.moh.gov.sg/docs/librariesprovider5/diseases-updates/weekly-infectious-disease-bulletin-year-2020ea2c0b1cec1549009844537d52f2377f.xlsx"
  #' 
  #' @param path The file path of the dataset.
  #' @return A table containing the combined historical epi-weekly records.
  #' @examples
  #' get_moh_eweekly("dataset.xlsx")
  
  # Read multiple sheets in an Excel Workbook
  dfs = path %>% 
    readxl::excel_sheets() %>% 
    set_names(., .) %>%
    lapply(readxl::read_xlsx, path = path, skip = 1)
  
  # Start and End date formats are different for 2020
  dfs[["2020"]]$Start = lubridate::dmy(dfs[["2020"]]$Start)
  dfs[["2020"]]$End = lubridate::dmy(dfs[["2020"]]$End)
  
  for (y in names(dfs)) {
    # Add a "Year" column to distinguish different tables
    dfs[[y]]$Year = y
    
    # There is some inconsistent naming of column headers
    names(dfs[[y]])[names(dfs[[y]]) %in% c("Campylobacter enterosis",
                                           "Campylobacterenterosis",
                                           "Campylobacteriosis")] <- "Campylobacter enteritis"
    names(dfs[[y]])[names(dfs[[y]]) == "Chikungunya Fever"] <- "Chikungunya"
    names(dfs[[y]])[names(dfs[[y]]) == "Dengue Haemorrhagic Fever"] <- "DHF"
    names(dfs[[y]])[names(dfs[[y]]) == "Dengue Fever"] <- "Dengue"
    names(dfs[[y]])[names(dfs[[y]]) %in% c("Hand, Foot and Mouth Disease",
                                           "Hand, Foot Mouth Disease")] <- "HFMD"
    names(dfs[[y]])[names(dfs[[y]]) == "Nipah virus infection"] <- "Nipah"
    names(dfs[[y]])[names(dfs[[y]]) == "Viral Hepatitis A"] <- "Acute Viral Hepatitis A"
    names(dfs[[y]])[names(dfs[[y]]) == "Viral Hepatitis E"] <- "Acute Viral Hepatitis E"
    names(dfs[[y]])[names(dfs[[y]]) %in% c("Zika Virus Infection",
                                           "Zika virus infection")] <- "Zika"
  }
  
  dplyr::bind_rows(dfs) %>% 
    dplyr::select(Year, everything()) %>% 
    dplyr::arrange(Year, `Epidemiology Wk`) %>% 
    dplyr::mutate(Year = as.numeric(Year))
}

get_mss_daily <- function(years, stations = c("Changi")) {
  #' Get Daily Records from Meteorological Service Singapore (MSS)
  #' 
  #' @description
  #' Available data ranges from Jan 1980 to May 2020.
  #' 
  #' This function will automatically calculate the epidemiological weeks.
  #' 
  #' @details http://www.weather.gov.sg/climate-historical-daily/
  #' 
  #' @param years A vector of years of interest.
  #' @param stations A vector of climate station names. Defaults to c("Changi").
  #' @return A table containing the combined historical daily records.
  #' @examples
  #' get_mss_daily(2014:2018)
  #' get_mss_daily(2012:2020, c("Changi", "Marine Parade", "Queenstown", "Sembawang"))
  
  dates = as.vector(t(outer(years, sprintf("%02d", 1:12), FUN = paste0)))
  
  # There's no data beyond May 2020 (at the moment), so remove the last 7
  #   months if 2020 is included
  if (2020 %in% years) {
    dates = dates[1:(length(dates) - 7)]
  }
  
  mappings = list(
    "Ang Mo Kio" = "109_",
    "Buangkok" = "55_",
    "Bukit Panjang" = "64_",
    "Changi" = "24_",
    "Jurong (West)" = "44_",
    "Khatib" = "122_",
    "Lower Peirce Reservoir" = "08_",
    "Marine Parade" = "113_",
    "Punggol" = "81_",
    "Queenstown" = "77_",
    "Seletar" = "25_",
    "Sembawang" = "80_",
    "Serangoon" = "36_",
    "Toa Payoh" = "88_",
    "Yishun" = "91_"
  )
  station_nums = sapply(stations, function(x) mappings[[x]])
  urls = paste0("http://www.weather.gov.sg/files/dailydata/DAILYDATA_S", 
                t(outer(station_nums, dates, FUN = paste0)), 
                ".csv")
  
  myreader = function(url) {
    tryCatch(
      readr::read_csv(url) %>% 
        # The "degree sign" in column headers are invalid multibyte characters
        # Data after Apr 2020 have slightly different column names
        # This would complicate joining the tables together
        # So, it might be better to manually label all column names
        setNames(., c("Station",
                      "Year",
                      "Month",
                      "Day",
                      "Daily_Rainfall_Total_mm",
                      "Highest_30_Min_Rainfall_mm",
                      "Highest_60_Min_Rainfall_mm",
                      "Highest_120_Min_Rainfall_mm",
                      "Mean_Temperature_degC",
                      "Maximum_Temperature_degC",
                      "Minimum_Temperature_degC",
                      "Mean_Wind_Speed_kmh",
                      "Max_Wind_Speed_kmh")) %>% 
        # The "long dashes" (em dash?) representing missing values are invalid
        #   multibyte characters and are replaced with NA (by coercion)
        dplyr::mutate_at(dplyr::vars(-Station), as.numeric),
      error = function(e) { NA }
    )
  }
  
  dfs = lapply(urls, myreader)
  
  dplyr::bind_rows(dfs[!is.na(dfs)]) %>% 
    dplyr::mutate(Date = paste(Year, Month, Day, sep = "-"),
                  Epiweek = lubridate::epiweek(Date)) %>% 
    dplyr::select(Station, Epiweek, everything(), -Date)
}

# w_data <- get_mss_daily(years = 2012:2020,
#                     stations = c("Changi",
#                                  "Marine Parade",
#                                  "Queenstown",
#                                  "Sembawang"))

weather <- readr::read_csv("../results/weather_daily_2012_2020.csv")
bulletin <- get_moh_eweekly("../data/weekly-infectious-disease-bulletin-year-2020ea2c0b1cec1549009844537d52f2377f.xlsx")

grouped_epiweek <- weather %>%
  dplyr::filter(Station == "Changi") %>% 
  dplyr::select(Year, Epiweek, matches("Rainfall_Total|Temperature"), -Station) %>% 
  dplyr::group_by(Year, Epiweek) %>% 
  dplyr::summarise(Rainfall_Total = sum(Daily_Rainfall_Total_mm),
                   Mean_Temperature = mean(Mean_Temperature_degC))

df <- bulletin %>% 
  dplyr::rename(Epiweek = `Epidemiology Wk`) %>% 
  dplyr::select(Year, Epiweek, Dengue) %>% 
  tidyr::drop_na() %>% 
  dplyr::left_join(grouped_epiweek, c("Year", "Epiweek")) %>% 
  tibble::add_column(x = 1:nrow(.))
```

```{r, explore_ncases_eweeks, fig.cap="", echo=FALSE, warning=FALSE, fig.width=8, fig.height=6}
f1 <- df %>% 
  ggplot(aes(x = x, y = Dengue)) + 
  geom_line(size = 1, color = "tomato3") + 
  scale_x_continuous(breaks = c(1 + 52 * (0:8)),
                     label = c(2012:2020)) + 
  labs(title = "Number of weekly Dengue cases of over time from 2012 to 2020",
       subtitle = "Why are the numbers abnormally high this year?",
       x = "",
       y = "Number of Cases",
       caption = "Source: moh.gov.sg")

f2 <- df %>% 
  ggplot(aes(x = x, y = Rainfall_Total)) + 
  geom_line(size = 1, color = "deepskyblue4") + 
  scale_x_continuous(breaks = c(1 + 52 * (0:8)),
                     label = c(2012:2020)) + 
  labs(title = "Total weekly Rainfall from 2012 to 2020",
       subtitle = "",
       x = "",
       y = "Total Rainfall (mm)",
       caption = "Source: weather.gov.sg")

f3 <- df %>% 
  ggplot(aes(x = x, y = Mean_Temperature)) + 
  geom_line(size = 1, color = "darkgreen") + 
  scale_x_continuous(breaks = c(1 + 52 * (0:8)),
                     label = c(2012:2020)) + 
  labs(title = "Average weekly Temperature from 2012 to 2020",
       subtitle = "",
       x = "",
       y = "Temperature (\u00b0C)",
       caption = "Source: weather.gov.sg")

gridExtra::grid.arrange(f1, f2, f3, nrow = 3, top = "Is there a relationship?")
```



## Overview of the Project {data-background=https://research.a-star.edu.sg/cms/figure/index/55e43a55140ba0db748b4b69.jpg data-background-size=auto data-background-position="right top"}
### Questions
  - Does atmospheric variables influence the incidence of dengue cases?
    - Specifically, is higher humidity, precipitation, or temperature (X) associated with increased numbers of dengue cases (Y)?
  - Can atmospheric variables predict the number of dengue cases?
  - Does the number of dengue cases increase with the number of COVID-19 cases?
  - Explain the different number of cases in different regions of Singapore.



## Data
- Daily climate records from [Meteorological Service Singapore (MSS)](http://www.weather.gov.sg/climate-historical-daily/)
  - Relevant variables: Rainfall, Temperature
  - Data for 2012 to 2020 will be gathered from selected climate stations
- Weekly case numbers (2012-W01 to 2020-W25) from [Ministry of Health (MOH)](https://www.moh.gov.sg/docs/librariesprovider5/diseases-updates/weekly-infectious-disease-bulletin-year-2020ea2c0b1cec1549009844537d52f2377f.xlsx)
- Latest number of cases by named regions from [National Environment Agency (NEA)](https://www.nea.gov.sg/dengue-zika/dengue/dengue-clusters)
- [Yearly population distribution across named regions](../data/respopagesextod2011to2019.csv)
- COVID-19 cases (Apr - Jul 2020)
  - ?

## Deprecated Data
- Monthly Air Temperature And Sunshine, Relative Humidity And Rainfall from [Singapore Department of Statistics (DOS)](https://www.tablebuilder.singstat.gov.sg/publicfacing/api/csv/title/15306.csv)
  - <u>Reason</u>: Higher resolution (daily) data available from MSS
  - Might reconsider if humidity data is needed
- Latest number of cases by regions with geocoordinates from [Data.gov.sg](https://data.gov.sg/dataset?q=Dengue)
  - <u>Reason</u>: Potentially **grandiose** (we've not learned leaflet yet!)

### Discarded
- Weekly case numbers (2012-W01 to 2020-W20) from [Data.gov.sg](https://data.gov.sg/dataset?q=Dengue)
  - <u>Reason</u>: More up-to-date data available from MOH



## Analysis Plan
- Transform dengue cases data
  - Converting epidemiology weeks cases to monthly cases for 2012 - 2020
  - Combine the different Excel sheets into a table 
- Compare the monthly cases across the years
  - Using repeated-measures ANOVA
  - Plotting
- To understand if atmospheric variables have an influence in the incidence of dengue cases
  - Correlate the weather variables with number of dengue cases monthly across the years
- To test if atmospheric variables predict number of dengue cases
  - Create models for regression
  - Compare the predicted value with actual value

## Analysis Plan
- Explain the current trend in differences between the number of cases in different regions
  - According to Strait's Times report, East region has the largest number of clusters than North (and West?) regions
  - Pick Marine Parade and Sembawang (or Queenstown?) to find out if there is a difference in their weather variables

### Considerations and limitations
- Time lag: Dengue cases manifest 1-2 weeks after infection
  - So, adjust timings for different datasets accordingly
- Seasonal effects
  - "Vector" months (June, July, August, September, October)
- COVID-19 cases vs. dengue cases
  - 90% foreign workers -> expect correlation
  - Number of cases from Apr - Jul 2020



# Preliminary Survey

```{r pre_explore, echo=FALSE, include=FALSE}
grouped_month <- weather %>%
  dplyr::select(Station, Year, Month, matches("Rainfall_Total|Temperature")) %>% 
  dplyr::group_by(Station, Year, Month) %>% 
  dplyr::summarise(Rainfall_Total = sum(Daily_Rainfall_Total_mm),
                   Mean_Temperature = mean(Mean_Temperature_degC),
                   Maximum_Temperature = mean(Maximum_Temperature_degC),
                   Minimum_Temperature = mean(Minimum_Temperature_degC))
```

---

```{r, explore_weather, echo=FALSE, fig.cap="", fig.width=8, fig.height=5}
w1 <- grouped_month %>% 
  dplyr::filter(Station == "Changi") %>% 
  tibble::add_column(x = 1:nrow(.)) %>% 
  dplyr::select(-Rainfall_Total) %>% 
  tidyr::gather("Variable",
                "Value",
                Mean_Temperature,
                Maximum_Temperature,
                Minimum_Temperature) %>% 
  ggplot(aes(x = x, y = Value, color = Variable)) + 
  scale_x_continuous(breaks = c(1 + 12 * (0:8)),
                     label = c(2012:2020)) +
  geom_line(size = 1) + 
  labs(title = "Average monthly Temperature from 2012 to 2020",
       subtitle = "",
       x = "",
       y = "Temperature (\u00b0C)",
       caption = "Source: weather.gov.sg")

w2 <- grouped_month %>% 
  dplyr::select(Station, Year, Month, Rainfall_Total) %>% 
  dplyr::mutate(x = (Year - 2012) * 12 + Month) %>% 
  ggplot(aes(x = x, y = Rainfall_Total, color = Station)) + 
  scale_x_continuous(breaks = c(1 + 12 * (0:8)),
                     label = c(2012:2020)) +
  geom_line(size = 1, alpha = 0.5) + 
  labs(title = "Total monthly Rainfall from 2012 to 2020",
       subtitle = "",
       x = "",
       y = "Total Rainfall (mm)",
       caption = "Source: weather.gov.sg")

gridExtra::grid.arrange(w1, w2, nrow = 2)
```
